\chapter{Theoretische und technologische Grundlagen}

\section{Künstliche Intelligenz und Machine Learning}
\subsection{Grundbegriffe}
\subsection{Überwachungs- und Analyseverfahren}
\subsection{Empfehlungssysteme und Predictive Analytics}

\section{Digitale Informationssysteme}
\subsection{Systemarchitekturen}
\subsection{Datenflüsse und Entscheidungslogiken}
\subsection{Plattformökosysteme}

\section{Monitoring- und Bewertungssysteme}
\subsection{Datenerfassung}
\subsection{Analyse und Modellbildung}
\subsection{Algorithmische Bewertung und Klassifikation}

\section{Methodisches Vorgehen}

\subsection{Forschungsansatz}
This thesis follows a literature-based research approach with a qualitative and interdisciplinary orientation. The objective of the thesis is to examine how AI-supported monitoring systems function as digital information systems and to analyze how these systems enable data-driven behavioral analysis and influence within digital environments.

Due to the socio-technical nature of the topic, the research combines perspectives from business informatics, information systems research, artificial intelligence, digital platform economics, and social sciences. The focus is placed on the technological structures, data processing mechanisms, and socio-economic implications of AI-supported monitoring and recommendation systems.

The thesis does not aim to empirically measure user behavior through experiments or surveys. Instead, existing scientific literature, technical publications, and selected practical examples are analyzed and critically compared.

\subsection{Literaturrecherche}
The literature research process was conducted iteratively and thematically. At the beginning of the research process, exploratory searches were performed in order to identify relevant concepts, technological mechanisms, and central authors within the research domains of artificial intelligence, digital monitoring systems, recommendation algorithms, platform ecosystems, and behavioral influence.

The identification of relevant literature was supported by AI-assisted exploratory search methods as well as academic search platforms and scientific databases. During the research process, scientific publications were continuously expanded through citation tracking and thematic cross-referencing. In particular, references and citations within already identified scientific papers were used to locate additional relevant literature and foundational works.

The literature selection focused primarily on peer-reviewed journal articles, conference papers, scientific books, and academically reputable publications within the fields of Information Systems, Artificial Intelligence, platform economics, and digital governance.

\subsection{Auswahlkriterien der Quellen}
The selection of literature was based on several quality criteria. Preference was given to scientific and peer-reviewed publications with thematic relevance to AI-supported monitoring systems, digital information systems, algorithmic decision-making, recommendation systems, surveillance mechanisms, and platform-based behavioral analysis.

In addition to scientific literature, selected reports, documentaries, and journalistic sources were included in limited form where they contributed practical examples or illustrated current societal debates. Such sources were not treated as scientific evidence but rather as supplementary contextual material.

Particular attention was paid to the scientific reputation of authors, publication quality, citation frequency, topical relevance, and methodological transparency of the respective sources.

\subsection{Methodische Grenzen}
The research topic contains several methodological limitations. Many modern AI-supported monitoring and recommendation systems operate as proprietary systems, meaning that internal algorithms, data structures, and optimization mechanisms are often not fully transparent to external researchers.

Furthermore, discussions surrounding digital surveillance, platform governance, and state-supported monitoring systems are frequently influenced by political, ideological, and media-related perspectives. As a result, the available literature and public information may contain normative interpretations and differing viewpoints.
